---
title: "Lodgepole Experiment"
author: "Olamide Adu"
date: "2022-11-02"
output:
  md_document:
    variant: markdown_github
---

# LodgePole Pine - Experiment (VII) (2)
This experiment is a soil scarification experiment with Lodegepole Pine in Västerbotten, Sweden. The treatments were either planting after soil scarification using deep plowing (deep), or planting without any site preparation (control). The size of the treatment plots were 30*30 m and the experiment was replicated in 4 blokcs. The plantation was made in 1988 and the experiment was remeasured after several years, to 2012, when all trees in the treatment plots were cross calipered at dbh (mm).

## Questions
For this data,
I'll be exploring the data and then calculating the basal area/ha and stem density/ha for every block and treatment. I seek to answer the following:

- What effect have a radical soil scarification method such as deep plowing on:
    
    - survival and
        
    - growth of lodgepole pine on the site

In other words: Is there a significant effect of treatment  if we test for:


-   stem density and

-   basal area.

```{r results='hide', message=FALSE, warning=FALSE}
# loading libraries

library(doBy)
library(dplyr)
library(lattice)
library(ggplot2)
library(car)
library(data.table)
library(TukeyC)
```

**Importing the data**
```{r}
L_pine <- read.table('https://raw.githubusercontent.com/xrander/Slu_experiment/master/Data/Lab%207/deepcult_lodgepolepine.txt',
           header = T,
           sep = '\t',
           na.strings = 'NA',
           dec = '.',
           strip.white = T)
head(L_pine)
```
**Little Data Exploration**
```{r}
str(L_pine)
```
I will change the data types of 'plot', 'block', and 'treatment' into factor data type

```{r}
plot(L_pine$dbh1, L_pine$dbh2,
     col = 'purple',
     xlab = 'dbh1',
     ylab = 'dbh2',
     main = 'DBH vs DBH')
```

```{r}
L_pine$plot <- as.factor(L_pine$plot)
L_pine$block <- as.factor(L_pine$block)
L_pine$treatment <- as.factor(L_pine$treatment)
```

### Survival
```{r}
with(L_pine, table(treatment, dead))
```
**subsetting only living trees**
```{r}
L_pine$dead <- ifelse(is.na(L_pine$dead) == TRUE,0 , L_pine$dead)
```

```{r}
L_pine1 <- subset(L_pine, dead== 0) # this returns values for only living trees, 1 means dead and 0 means alive
```


```{r}
summary(L_pine1)
```

```{r}
bwplot(dbh1~plot | treatment, data = L_pine1,
       main = 'Dbh across plots for Treatments',
       ylab = 'dbh(cm)',
       xlab = 'plots')
```

**getting the mean dbh**
```{r}
L_pine1$dbh <- (L_pine1$dbh1 + L_pine1$dbh2)/2000
```

**Evaluating basal area**
```{r}
L_pine1$ba <- (L_pine1$dbh/2)^2 * pi
```

```{r}
summary(L_pine1)
```

**Basal area of only living trees**
```{r}
L_pineba <- summaryBy (ba + dbh ~plot+block+treatment,
data = L_pine1,
na.rm = T,
keep.names = T,
FUN = sum)

head(L_pineba)
```
**Estimating basal area per hectare**
```{r}
L_pineba$baha <- L_pineba$ba * 10000/(30*30)
```

**Visualizing basal area per hectare**

```{r}
barchart(baha~block,
         group = treatment,
         data = L_pineba,
         ylab = 'basal area per hectare (m3 ha-1)',
         xlab = 'plots',
         ylim = c(0, 30),
         main = 'Basal area per hectare across plots',
         auto.key = list(corners = c(0.05, 0.95)))
```

### Stem Density
```{r}
L_pinedens <- summaryBy(ba~plot+block+treatment,
                        data = L_pine1, FUN = length)
L_pinedens$dens_ha <- L_pinedens$ba.length * 10000 / (30*30) 
```

```{r}
barchart(dens_ha~block,
         group = treatment,
         data = L_pinedens,
         ylab = 'Density per Hectare',
         xlab = 'plots',
         ylim = c(0, 3000),
         main = 'Density of Stand',
         auto.key = list(corner = c(0.1, 0.95)))
```

## Analysis of Variance
```{r}
M_ba <- lm(baha~block+treatment,
   data = L_pineba)
Anova(M_ba)
```

```{r}
M_dens <- lm(dens_ha~block+treatment,
   data = L_pinedens)
Anova(M_dens)
```
Basal area was significantly higher in the deep plowing treatment but has no significant difference in stem density indicating that the primary reason to the good response of the deep plowing has brrn on growth more than on survival.

[Previous page](pct.md)
<br>

[Homepage](Readme.md)
<br>

[Next Page](growth_experiment.md)
<br>

[Back to portfolio](https://xrander.github.io)